---
title: "Bad data"
author: "Ben Williams"
date: "July 15, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message = FALSE)
```

A "problem" with reproducible research is that you need to keep all of your error corrections in code.

```{r}
library(tidyverse)
library(readxl)
library(lubridate)
```

First import the data
```{r}
read_xlsx("bad_good_data.xlsx", sheet=1)
```

This didn't work out so well, because the column names are multi-line.

Remove the column names on import and name the data "bad". 

```{r}
bad <- read_xlsx("bad_good_data.xlsx", sheet=1, range='a5:w150', col_names=F)
```

Take a look at bad.

```{r}
glimpse(bad)

```

Right off the bat we can see a problem - The first column "year" is a character value.
```{r}
glimpse(bad$X__1)

bad %>% 
  mutate(year = ifelse(is.character(X__1), 20180, X__1)) %>% glimpse()

bad %>% 
  mutate(year = as.numeric(gsub("\\*", "\\", X__1)),
         year = ifelse(year < 1999, year + 2000, year)) -> bad
bad$year



```
Drop any column without data, or with data estimated in an unknown fashion.

```{r}
less_bad = bad[,c(1:5, 9:12, 13, 14, 16, 18, 20)]
```

Add names to the columns
```{r}
names(less_bad) <- c('year', 'month','date', 'gear_type', 'Mgmt_area', 'Sample_type', 'stat_area', 'age', 'Sex', 'length_cm', 'weight_kg', 'maturity', 'lat', 'long')
```

Check the new data set

```{r}
glimpse(less_bad)
```

Now it is time for cleanup...

Go through each column and double check the unique values. Correct any issues. 

```{r}
sapply(less_bad, unique)
```

 - We can see that the years after 2010 are incorrect
 - Therefore the dates have NA (likely due to a function being used in Excel)
 - There is a lower and upper case 'CSEO' management area
 - Sex has multiple formats
 - length_cm has a number of entries that look to be in mm
 - weight_kg has a number of entries that look to be in grams
 - multiple maturity levels (numeric and text) are present

```{r}
less_bad %>% 
  mutate(Mgmt_area = ifelse(Mgmt_area=='cseo', 'CSEO', Mgmt_area),
         Sample_type = 'Random',
         Sex = case_when(Sex=='1' ~ 'male',
                         Sex=='m' ~ 'male',
                         Sex=='2' ~ 'female',
                         Sex=='f' ~ 'female',
                         Sex=='F' ~ 'female'),
         maturity = case_when(maturity=='immature' ~ 1,
                              maturity=='developing' ~ 2,
                              maturity=='spwning'|maturity=='spawning' ~ 3,
                              maturity=='spent' ~ 4,
                              maturity=='1' ~ 1,
                              maturity=='2' ~ 1,
                              maturity=='3' ~ 2,
                              maturity=='4' ~ 3,
                              maturity=='5' ~ 3,
                              maturity=='6' ~ 4,
                              maturity=='7' ~ 5)) -> less_bad

glimpse(less_bad)
```

Examine the age, length data, etc.

```{r}
ggplot(less_bad, aes(age, length_cm)) + geom_point()
```

```{r}
ggplot(less_bad, aes(length_cm)) + geom_density(fill = 4, alpha = .3)
min(less_bad$length_cm)
max(less_bad$length_cm)
less_bad %>% 
  mutate(length_cm = ifelse(length_cm > 75, length_cm / 100, length_cm)) %>% 
  ggplot(aes(length_cm)) + geom_density(fill = 4, alpha = .3)
```

Same thing for weight

```{r}

ggplot(less_bad, aes(weight_kg)) + geom_density(fill = 4, alpha = .3)

less_bad %>% 
  mutate(weight_kg = ifelse(weight_kg > 7.5, weight_kg / 100, weight_kg)) %>% 
  ggplot(aes(weight_kg)) + geom_density(fill = 4, alpha = .3)
```

This should work, however it throws and error of a "non-finite value". Look at cell N135 in the spreadsheet. This isn't an accepted "NA" value so it causes a problem. 

```{r}
less_bad %>% 
  mutate(weight_kg = ifelse(is.numeric(weight_kg), weight_kg, NA),
        weight_kg = ifelse(weight_kg > 7.5, weight_kg / 100, weight_kg)) %>% 
  ggplot(aes(weight_kg)) + geom_density(fill = 4, alpha = .3)
```

Double check ages

```{r}
ggplot(less_bad, aes(age)) + geom_density(fill = 4, alpha = .3)
```

```{r}
less_bad %>% 
  mutate(length_cm = ifelse(length_cm > 75, length_cm / 100, length_cm),
         weight_kg = ifelse(is.numeric(weight_kg), weight_kg, NA),
        weight_kg = ifelse(weight_kg > 7.5, weight_kg / 100, weight_kg)) -> better 
```
